---
title: "p8105_hw1_mjd2283"
author: "Max"
date: "2026-09-23"
output: github_document
---

## Problem 1

```{r penguins}
library(tidyverse)

data("penguins", package = "palmerpenguins")
summary(penguins)
```
In this dataset, the names of species include `r levels(pull(penguins,species))`, and the names of islands include `r levels(pull(penguins,island))`.

Sex is either (if not NA): `r levels(pull(penguins,sex))`.

The median length of bills is `r median(pull(penguins,bill_length_mm), na.rm = TRUE)` mm.

The median length of flippers is `r median(pull(penguins,flipper_length_mm), na.rm = TRUE)` mm.

The median body mass weight is `r median(pull(penguins,body_mass_g), na.rm = TRUE)` g.

The min year is `r min(pull(penguins,year))` and the max is `r max(pull(penguins,year))`.

Used AI to show how to use r in line code to show names/values. R levels gets the categories of factors in the dataset. For the numeric values, I used median values besides year. I reported the mean flipper length below as well. 

There are `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

na.rm is a logical evaluating to TRUE or FALSE indicating whether NA values should be stripped before the computation proceeds. Used AI to figure out this argument usage

```{r}
## scatterplot of bill length (X) and flipper length (y)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_plot.pdf", height = 4, width = 6)
```

Used ggplot to do the scatter plot and add color, and then saved a pdf using ggsave. 

## Problem 2

```{r}

set.seed(1)

problem2_df = tibble(
  vec_numeric= rnorm(10),
  vec_logical= c(vec_numeric>TRUE),
  vec_character= c("It", "is", "cold", "outside", "but", "I", "am", "inside", "right", "now"),
  vec_factor = c("fast", "slow", "fast", "fast", "medium", "fast", "fast", "medium", "slow", "slow")
)

print.data.frame(problem2_df)

library(tidyverse)
mean(pull(problem2_df,vec_numeric))
mean(pull(problem2_df,vec_logical))
mean(pull(problem2_df,vec_character))
mean(pull(problem2_df,vec_factor))

```
I already loaded tidyverse in the first coding section. But put it above again since HW mentioned it

Used print.data.frame since I wanted to see the results.


```{r, results = "hide"}
as.numeric(pull(problem2_df,vec_logical))
as.numeric(pull(problem2_df,vec_character))
as.numeric(pull(problem2_df,vec_factor))
```

For the vec_logical, it gives 0 and 1, so the statement is either False (0) or True (1).

For the other ones vec_character and vec_factor, these are NA since these values can't turn into numbers. 











