---
title: "p8105_hw1_mjd2283"
author: "Max"
date: "2026-09-23"
output: github_document
---

```{r penguins}
data("penguins", package = "palmerpenguins")
summary(penguins)
nrow(penguins)
ncol(penguins)
``` 
```{r}
library(tidyverse)
mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```
In this dataset, the names of categories include `r levels(pull(penguins,species))`, and  `r levels(pull(penguins,island))`.

The min length of bills (mm) is `r min(pull(penguins,bill_length_mm))`, and the max is `r max(pull(penguins,bill_length_mm))`



`r levels(pull(penguins,bill_depth_mm))`, `r levels(pull(penguins,flipper_length_mm)`, `r levels(pull(penguins,body_mass_g))`, `r levels(pull(penguins,sex))`, `r levels(pull(penguins,year))`

Used AI to show how to use r in line code to show names/values. r levels gets the factos in the dataset. For the numeric values, I used min and max. 

There are `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

na.rm is a logical evaluating to TRUE or FALSE indicating whether NA values should be stripped before the computation proceeds. Used AI to figure out this argument usage

## Problem 1

```{r}
## scatterplot of x and y
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

getwd()
ggsave("penguins_plot.pdf", height = 4, width = 6)
```

Used ggplot to do the scatter plot and add color, and then saved a pdf using ggsave. 

## Problem 2


```{r}

set.seed(1)

problem2_df = tibble(
  vec_numeric= rnorm(10),
  vec_logical= c(vec_numeric>TRUE),
  vec_character= c("It", "is", "cold", "outside", "but", "I", "am", "inside", "right", "now"),
  vec_factor = c("fast", "slow", "fast", "fast", "medium", "fast", "fast", "medium", "slow", "slow")
)

print.data.frame(problem2_df)

library(tidyverse)
mean(pull(problem2_df,vec_numeric))
mean(pull(problem2_df,vec_logical))
mean(pull(problem2_df,vec_character))
mean(pull(problem2_df,vec_factor))

```
I already loaded tidyverse in the first coding section. But put it above again since HW mentioned it

Used print.data.frame since I wanted to see the results.


```{r, results = "hide"}
as.numeric(pull(problem2_df,vec_logical))
as.numeric(pull(problem2_df,vec_character))
as.numeric(pull(problem2_df,vec_factor))
```

For the vec_logical, it gives 0 and 1, so the statement is either False (0) or True (1).

For the other ones vec_character and vec_factor, these are NA since these values can't turn into numbers. 











